Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-02-27 Manuscript/manuscript.Rmd
added more info for AUPRC; commented out other metrics in case we decide to use them

Commonly used measures of model performance for binary classifiers include threshold-free metrics like the area under the receiver operating curve (AUROC), and other threshold-specific metrics like accuracy, sensitivity, and specificity. The ROC curve is created by plotting the true positive rate (TPR) against the false positive rate (FPR) at various threshold settings. The subsequent AUROC, measures how well the model is capable of distinguishing between the two classes [@park2004receiver]. Accuracy is defined as the ratio of correctly identified cases (either from the positive or negative class) over the total number of cases. It measures the overall proportion of correctly classified instances. Sensitivity, also referred to as recall or TPR, is the ratio of true positives over the total number of positive classes. Specificity, on the other hand, also referred to as TNR, is the ratio of true negatives over the total number of negative classes. Each measures the proportion of correctly identified positive and negative classes respectively [@yerushalmy1947statistical]. Problems arise when using such measures in the event of imbalanced classes. For AUROC, when the prevalence of an event is low--that is, a defined genomic region truly containing a TAD boundary--mean probabilities are biased towards the majority class. This results in the area of interest under an ROC curve to be compressed toward a small corner in the ROC space [@prati2011survey]. While not directly affecting the value of the AUROC, this can have severe consequences on the interpretability of the model's performance. A better threshold-free measure to use when dealing with imbalanced data is the area under the precision-recall curve (AUPRC).  Likewise, threshold-specific metrics like accuracy and specificity are also problematic when evaluating model performance on imbalance data due to being highly influenced by the correct classification of the majority class. Certain composite threshold-specific metrics such as balanced accuracy and Youden's Index should be preferred because they soften the influence of a model's ability to correctly classify majority classes and yield more dependable and interpretable results [@jeni2013facing]. Here, balanced accuracy measures the average accuracy obtained on either class [@brodersen2010balanced]. Informedness, also referred to as Youden's Index, measures the probability of the model making an informed decision and corresponds to giving equal weight per class, rather than to each instance like Accuracy [@youden1950index;@powers2011evaluation].

>>>>

Commonly used measures of model performance for binary classifiers include threshold-free metrics like the area under the receiver operating curve (AUROC), and other threshold-specific metrics like accuracy, sensitivity, and specificity. The ROC curve is created by plotting the true positive rate (TPR) against the false positive rate (FPR) at various threshold settings. The subsequent AUROC, measures how well the model is capable of distinguishing between the two classes [@park2004receiver]. Accuracy is defined as the ratio of correctly identified cases (either from the positive or negative class) over the total number of cases. It measures the overall proportion of correctly classified instances. Sensitivity, also referred to as recall or TPR, is the ratio of true positives over the total number of positive classes. Specificity, on the other hand, also referred to as TNR, is the ratio of true negatives over the total number of negative classes. Each measures the proportion of correctly identified positive and negative classes respectively [@yerushalmy1947statistical]. Problems arise when using such measures in the event of imbalanced classes. For AUROC, when the prevalence of an event is low--that is, a defined genomic region truly containing a TAD boundary--mean probabilities are biased towards the majority class. This results in the area of interest under an ROC curve to be compressed toward a small corner in the ROC space [@prati2011survey]. While not directly affecting the value of the AUROC, this can have severe consequences on the interpretability of the model's performance. A better threshold-free measure to use when dealing with imbalanced data is the area under the precision-recall curve (AUPRC).  The interpretation of the AUPRC of a model is similar to that of the AUROC. However, where as the baseline model performance according to the AUROC is given by 0.50, the baseline model performance according to the AUPRC is given by the ratio of true positives over the total number of cases. This enables a more realistic benchmark to evaluate a model on as this ratio will be lower than 0.5 for a test data set with imbalanced classes. Likewise, threshold-specific metrics like accuracy and specificity are also problematic when evaluating model performance on imbalance data due to being highly influenced by the correct classification of the majority class. Certain composite threshold-specific metrics such as balanced accuracy and Youden's Index should be preferred because they soften the influence of a model's ability to correctly classify majority classes and yield more dependable and interpretable results [@jeni2013facing]. Here, balanced accuracy measures the average accuracy obtained on either class [@brodersen2010balanced]. Informedness, also referred to as Youden's Index, measures the probability of the model making an informed decision and corresponds to giving equal weight per class, rather than to each instance like Accuracy [@youden1950index;@powers2011evaluation].

<!--
Better measures to consider are composite metrics involving both precision and recall together. Therefore, measures like the F1-Score, Youden's Index (YI), and Matthew's Correlation Coefficient (MCC) offer better interpretation, especially in the instance of imbalanced classes [@jeni2013facing]. The F1-Score can be interpreted as a weighted average of the precision and recall values, while Youden's index is interpreted as the arithmetic mean between sensitivity (recall) and specificity. MCC is, in essence, the correlation coefficient between the observed and predicted binary classifications. Each metric uses all 4 quadrants of the confusion matrix, thus providing more balanced measures in the case of imbalanced data [@chicco2017ten].
-->
